---
title: "Business Intelligence Lab Submission Markdown"
author: "<>"
date: "<23/10/2023>"
output:
  github_document: 
    toc: yes
    toc_depth: 4
    fig_width: 6
    fig_height: 4
    df_print: default
editor_options:
  chunk_output_type: console
---

# Student Details

+---------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **Student ID Numbers and Names of Group Members** | *\<list one Student name, class group (just the letter; A, B, or C), and ID per line, e.g., 123456 - A - John Leposo; you should be between 2 and 5 members per group\>* |
|                                                   |                                                                                                                                                                          |
|                                                   | 1.  128998 - B - Crispus Nzano                                                                                                                           |
|                                                   |                                                                                                                                                                      
+---------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **GitHub Classroom Group Name**                   | BI-Loan-Appraisal-Project                                                                                                       |
+---------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **Course Code**                                   | BBT4206                                                                                                                                                                  |
+---------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **Course Name**                                   | Business Intelligence II                                                                                                                                                 |
+---------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **Program**                                       | Bachelor of Business Information Technology                                                                                                                              |
+---------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **Semester Duration**                             | 21^st^ August 2023 to 28^th^ November 2023                                                                                                                               |
+---------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------+

# Setup Chunk

We start by installing all the required packages, each Issue and Milestone will have its own packages.

```{r Install Packages, echo=TRUE, message=FALSE, warning=FALSE}
## formatR - Required to format R code in the markdown ----

if (require("languageserver")) {
  require("languageserver")
} else {
  install.packages("languageserver", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

# Introduction ----
# Resampling methods are techniques that can be used to improve the performance
# and reliability of machine learning algorithms. They work by creating
# multiple training sets from the original training set. The model is then
# trained on each training set, and the results are averaged. This helps to
# reduce overfitting and improve the model's generalization performance.

# Resampling methods include:
## Splitting the dataset into train and test sets ----
## Bootstrapping (sampling with replacement) ----
## Basic k-fold cross validation ----
## Repeated cross validation ----
## Leave One Out Cross-Validation (LOOCV) ----

# STEP 1. Install and Load the Required Packages ----
## mlbench ----
if (require("mlbench")) {
  require("mlbench")
} else {
  install.packages("mlbench", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

## caret ----
if (require("caret")) {
  require("caret")
} else {
  install.packages("caret", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

## kernlab ----
if (require("kernlab")) {
  require("kernlab")
} else {
  install.packages("kernlab", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

## randomForest ----
if (require("randomForest")) {
  require("randomForest")
} else {
  install.packages("randomForest", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

install.packages("formatR")


```

------------------------------------------------------------------------

**Note:** the following "*KnitR*" options have been set as the defaults in this markdown:\
`knitr::opts_chunk$set(echo = TRUE, warning = FALSE, eval = TRUE, collapse = FALSE, tidy.opts = list(width.cutoff = 80), tidy = TRUE)`.

More KnitR options are documented here <https://bookdown.org/yihui/rmarkdown-cookbook/chunk-options.html> and here <https://yihui.org/knitr/options/>.

```{r setup, echo=TRUE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(
	eval = TRUE,
	echo = TRUE,
	warning = FALSE,
	collapse = FALSE,
	tidy = TRUE
)
```

------------------------------------------------------------------------

**Note:** the following "*R Markdown*" options have been set as the defaults in this markdown:

> output:\
> \
> github_document:\
> toc: yes\
> toc_depth: 4\
> fig_width: 6\
> fig_height: 4\
> df_print: default\
> \
> editor_options:\
> chunk_output_type: console

## Milestone 6 out of 8
# Loading the Loan Status Train Imputed Dataset


Issue 6 Training the Model. 
```{r Train Model}
## 6. Training the Model ----

if (!is.element("languageserver", installed.packages()[, 1])) {
  install.packages("languageserver", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}
require("languageserver")

# Introduction ----
# The performance of the trained models can be compared visually. This is done
# to help you to identify and choose the top performing models.

# STEP 1. Install and Load the Required Packages ----
## mlbench ----
if (require("mlbench")) {
  require("mlbench")
} else {
  install.packages("mlbench", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

## caret ----
if (require("caret")) {
  require("caret")
} else {
  install.packages("caret", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

## kernlab ----
if (require("kernlab")) {
  require("kernlab")
} else {
  install.packages("kernlab", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

## randomForest ----
if (require("randomForest")) {
  require("randomForest")
} else {
  install.packages("randomForest", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}



## STEP 2. Load the Dataset ----
library(readr)
loans <- read_csv("C:/Users/Cris/github-classroom/BI-Loan-Appraisal-Project/data/loans_imputed.csv")
View(loans)

## STEP 3. The Resamples Function ----

# Analogy: We cannot compare apples with oranges; we compare apples with apples.

# The "resamples()"  function checks that the models are comparable and that
# they used the same training scheme ("train_control" configuration).
# To do this, after the models are trained, they are added to a list and we
# pass this list of models as an argument to the resamples() function in R.

## 3.a. Train the Models ----
# We train the following models, all of which are using 10-fold repeated cross
# validation with 3 repeats:
#   LDA
#   CART
#   KNN
#   SVM
#   Random Fores

train_control <- trainControl(method = "repeatedcv", number = 10, repeats = 3)

### LDA ----
set.seed(7)
loans_imputed_model_lda <- train(Status ~ ., data = loans,
                                 method = "lda", trControl = train_control)

### CART ----
set.seed(7)
loans_imputed_model_cart <- train(Status ~ ., data = loans,
                                  method = "rpart", trControl = train_control)

### KNN ----
set.seed(7)
loans_imputed_model_knn <- train(Status ~ ., data = loans,
                                 method = "knn", trControl = train_control)

### SVM ----
set.seed(7)
loans_imputed_model_svm <- train(Status ~ ., data = loans,
                                 method = "svmRadial", trControl = train_control)

### Random Forest ----
set.seed(7)
loans_imputed_model_rf <- train(Status ~ ., data = loans,
                                method = "rf", trControl = train_control)

## 3.b. Call the `resamples` Function ----
# We then create a list of the model results and pass the list as an argument
# to the `resamples` function.

results <- resamples(list(LDA = loans_imputed_model_lda, CART = loans_imputed_model_cart,
                          KNN = loans_imputed_model_knn, SVM = loans_imputed_model_svm,
                          RF = loans_imputed_model_rf))

# STEP 4. Display the Results ----
## 1. Table Summary ----
# This is the simplest comparison. It creates a table with one model per row
# and its corresponding evaluation metrics displayed per column.

summary(results)

## 2. Box and Whisker Plot ----
# This is useful for visually observing the spread of the estimated accuracies
# for different algorithms and how they relate.

scales <- list(x = list(relation = "free"), y = list(relation = "free"))
bwplot(results, scales = scales)

## 3. Dot Plots ----
# They show both the mean estimated accuracy as well as the 95% confidence
# interval (e.g. the range in which 95% of observed scores fell).

scales <- list(x = list(relation = "free"), y = list(relation = "free"))
dotplot(results, scales = scales)

## 4. Scatter Plot Matrix ----
# This is useful when considering whether the predictions from two
# different algorithms are correlated. If weakly correlated, then they are good
# candidates for being combined in an ensemble prediction.

splom(results)

## 5. Pairwise xyPlots ----
# You can zoom in on one pairwise comparison of the accuracy of trial-folds for
# two models using an xyplot.

# xyplot plots to compare models
xyplot(results, models = c("LDA", "SVM"))

# or
# xyplot plots to compare models
xyplot(results, models = c("SVM", "CART"))

## 6. Statistical Significance Tests ----
# This is used to calculate the significance of the differences between the
# metric distributions of the various models.

### Upper Diagonal ----
# The upper diagonal of the table shows the estimated difference between the
# distributions. If we think that LDA is the most accurate model from looking
# at the previous graphs, we can get an estimate of how much better it is than
# specific other models in terms of absolute accuracy.

### Lower Diagonal ----
# The lower diagonal contains p-values of the null hypothesis.
# The null hypothesis is a claim that "the distributions are the same".
# A lower p-value is better (more significant).

diffs <- diff(results)

summary(diffs)

#The model of choice will be 'LDA'.This is as a result of the model givong the highest accuracy 
#of 0.8137810 as compared to the other models (CART 0.7947), (KNN 0.6530), (SVM 0.8094)
# AND (RF 0.8050)


# Upload *the link* to "Lab-Submission-Markdown.md" (not .Rmd)
# markdown file hosted on Github (do not upload the .Rmd or .md markdown files)
# through the submission link provided on eLearning.
#install the same package version in their local
# machine during their initialization step.
# renv::snapshot() 

```


**etc.** as per the lab submission requirements. Be neat and communicate in a clear and logical manner.








